Smart Representations: {R}ationality and Evolution in a Richer Environment

### 2016-06-25 paper/rev1_paperBlinded.tex
paper revisions first round complete

As we noticed, meta-games factor in statistical properties of the environment. For particular empirical
purposes, one would consult a specific class of games $\mathcal{G}$ with appropriate,
empirically informed probabilities $P_{\mathcal{G}}$ in order to match the natural environment of a given
population. 
%But for our theoretical purposes we adopt another approach. 
For our theoretical purposes, let $\mathcal{G}$ be a set of
symmetric two-player games with two acts for a start. A game is then individuated solely by its payoff
function, i.e., by a quadruple of numbers $G=(a,b,c,d)$. As for the occurrence probability $P_{\mathcal{G}}(G)$ of
game $G$, we imagine that the values $a,b,c,d$ are i.i.d.~random variables sampled from the set $ \lbrace 0, \dots, 10 \rbrace$ according to uniform probability $P_{V}$. The probability distribution of games is thus the fourfold product of $P_{V}$, that is, $P_{\mathcal{G}}=P_{V}^{4}$. Using Monte Carlo simulations, we can then
approximate the values of equation~(\ref{eq:FittnessChoiceMechGamePairwise})
to construct meta-game payoffs. Results based on $100,000$ randomly sampled games are given in
table~\ref{tab:ExpectedFitness_4Types}.\footnote{Concretely, $100,000$ games were sampled
  repeatedly by choosing independently four integers between 0 and 10 uniformly at random. For
  each game, the action choices of all four choice mechanisms were determined and payoffs from
  all pairwise encounters recorded. The number
  in each cell of table~\ref{tab:ExpectedFitness_4Types} is the average payoff for the choice
  mechanism listed in the row when matched with the choice mechanism in the column.}

@@

Simulation results of the (discrete time) \emph{replicator dynamics}
\citep{TaylorJonker1978:Evolutionary-St} indeed show that random initial population
configurations are attracted to states with only three player types:
$(\text{reg}, \text{imp})$, $(\text{reg}, \text{prc})$ and $(\text{obj}, \text{prc})$. The
relative proportions of these depend on the initial shares in the population. This variability disappears if
we add a small mutation rate to the dynamics. Take a fixed, small mutation rate $\epsilon$ for
the probability that a player's subjective utility \emph{or} her subjective belief changes to another
utility or belief. The probability that a player's subjective representation randomly mutates into a
completely different representation with altogether different utility and belief would
then be $\epsilon^2$. With these assumptions about ``component-wise mutations'', numerical
simulations of the (discrete time) \emph{replicator mutator dynamics}
\citep{Nowak2006:Evolutionary-Dy} show that already for very small mutation rates almost all
initial population states converge to a single fixed point in which the majority of players have
regret-based utility. For instance, with $\epsilon = 0.001$, almost all initial populations are
attracted to a final distribution with proportions:

@@

%Obviously, $(\text{reg}, \text{imp})$,
%$(\text{obj}, \text{prc})$, and $(\text{reg}, \text{prc})$ players will be neutrally stable, as before. 
%and fitness-based selection will tend to weed out maxmin play and leave us with a population of
%arbitrary frequency of $(\text{reg}, \text{imp})$, $(\text{obj}, \text{prc})$, and
%$(\text{reg}, \text{prc})$ players. 
The result shows that there is support for the main conceptual
point that we wanted to make: 
%non-veridical representations \emph{can}, under
%specific circumstances, persist under evolutionary selection based on objective
%fitness. Moreover,
objective preference representations are not necessarily favored by
natural selection. In particular, under generous circumstances objective preferences are outperformed by non-veridical regret preferences, if agents may have imprecise beliefs.
This
tells us that the main conclusions drawn in the previous section based on the approximated meta-game of table~\ref{tab:ExpectedFitness_4Types} hold more generally 
for
arbitrary $2 \times 2$ symmetric games with i.i.d. sampled payoffs.

Even more in general, we can show that the same result obtains for any imprecise belief, not only for maximally uncertain players. 
Let the uncertainty held by a player be defined by a closed and convex set of probabilities $ [s,t] \subseteq \Delta(A) $ over the co-player's actions, where $s$ is the lower probability and $t$ is the upper probability of action $II$.
We can then prove the following proposition, which
is the equivalent of Proposition~\ref{proposition1} for any possible (not necessarily
maximal) degree of uncertainty $[s,t]$, with $s \neq t$. There is only one difference: we are
now going to require i.i.d. drawing of a \emph{continuous} random variable. This is due
to the fact that, for arbitrarily small intervals $[s,t]$, objective players $([s,t],\text{obj})$
and regret players $([s,t],\text{reg})$ can behave as holding a unique probability measure
(precise belief) if the underlying payoff space is not dense. The reason for this technical requirement
will become clearer from the proof.

\begin{proposition} \label{proposition2}
In the class of $2\times2$ symmetric games generated by i.i.d. drawing
of a continuous random variable that takes values in a real interval
$(\underline{r},\overline{r})$ with probability density $P_V$, for
any imprecise belief $[s,t]$, the only
evolutionarily stable state of a population with regret players $([s,t],\text{reg})$
and objective players $([s,t],\text{obj})$ is a monomorphic state of $([s,t],\text{reg})$
players.
\end{proposition}

@@

\subsection{More Representations}
\label{sec:more-types}

Beyond objective and regret-based utilities, other subjective preferences have been investigated, especially in behavioral economics and in evolutionary game theory. A
famous example is the \textit{altruistic} preference \citep[e.g.,][]{Beck76,BestGuth98}, summoned to
explain the possibility of altruistic behavior. At the other end of the spectrum, the \emph{competitive}
preference is located. The two subjective utilities are defined as follows.
\begin{enumerate}
\item \textit{altruistic} utility: for all $G \in \mathcal{G}$, $\text{alt}^G(a,a') = \pi^G(a,a') + \pi^G(a',a)$;\footnote{A
    more general formulation would be to define $ \alpha$-altruistic utility, for
    $\alpha \in [0,1]$,
    $ u^G_\alpha(a, a')=\pi^G(a,a') + \alpha \pi^G(a',a)$. Since we are not
    interested in the evolution of degrees of altruism, here we simply fix $ \alpha = 1 $. Analogously for $\alpha$-competitive utilities too.}
\item \textit{competitive} utility: for all $G \in \mathcal{G}$, $\text{com}^G(a,a') = \pi^G(a,a') - \pi^G(a',a)$.
\end{enumerate}

@@

Table~\ref{tab:ExpectedFitness_2x2_Full} shows results of Monte Carlo simulations that
approximate the expected fitness in the relevant meta-game with all the subjective representations considered so
far. These results confirm basic intuitions about altruistic and competitive types: everybody would like to have an altruistic co-player and nobody
would like to play against a competitive player. Perhaps more surprisingly, $(\text{alt}, \text{imp})$
come up strictly dominated by $(\text{com}, \text{imp})$, but competitive types themselves are worse off
against all types except against maxmin players $(\text{obj}, \text{imp})$ than any of
the behaviorally equivalent types $(\text{reg}, \text{imp})$, $(\text{obj}, \text{prc})$, and
$(\text{reg}, \text{prc})$.
\iffalse This is a noteworthy results in the light of the fact
that evolving altruistic preferences have been shown to support cooperative behavior in a
single stage game \myalert{[CITE]}. In contrast, averaging over payoffs in multiple stage
games, like we do here, makes altruistic preferences prime victims of evolutionary eradication.
\fi
It is easy to see then that the previous results still obtain for the larger meta-game in
table~\ref{tab:ExpectedFitness_2x2_Full}: $(\text{reg}, \text{imp})$,
$(\text{obj}, \text{prc})$, and $(\text{reg}, \text{prc})$ are still neutrally stable;
simulation runs of the (discrete-time) replicator dynamics on the $8 \times 8$ meta-game from
table~\ref{tab:ExpectedFitness_2x2_Full} end up in population states consisting of only these
three types in arbitrary proportion.

@@

Results from Section~\ref{sec:basic-model-1} relied heavily on Fact~\ref{fact:equivalence2x2}, that
is no longer true when we look at arbitrary $n \times n$ games. 
Table~\ref{tab:ExpectedFitness_10x10} gives approximations of expected fitness in the class of
$n \times n$ symmetric games. Concretely,
the numbers in table~\ref{tab:ExpectedFitness_10x10} are averages of evolutionary payoffs obtained in
100,000 randomly sampled symmetric games, where each game was sampled by first picking a number of acts
$n \in \set{2, \dots, 10}$ uniformly at random, and then filling the necessary $n \times n$
payoff matrix with i.i.d.~sampled numbers, as before.

@@

On the other hand, there are plenty of ways in which the basic insight from
Propositions~\ref{proposition1} and \ref{proposition2} that regret-based utilities are strictly better than
veridical objective utilities in case of imprecise beliefs can make for situations in which
evolution would select for regret types exclusively, even in $n \times n$ games. If, for
example, the belief of a player is a trait that biological evolution has no bite on, but rather
something that the particular choice situation would exogenously give us (possibly because of the different amount of information available in different choice situations), then regret-based
preferences can again drive out veridical preferences altogether. For example,
suppose that only preferences compete and that agents' beliefs are exogenously given,
in such a way that both players hold precise (Bayesian) uniform beliefs with probability $p$ and they both have maximally imprecise
beliefs otherwise. This transforms the meta-game from table~\ref{tab:ExpectedFitness_10x10} into a
simpler $4 \times 4$ meta-game in which the payoff obtained by a subjective utility is the weighted
average over the payoffs of the subjective representations including that utility in
table~\ref{tab:ExpectedFitness_10x10}. Setting $p = 0.98$ for illustration, we get the meta-game in table~\ref{tab:ExogeneousEpistemics}.
\begin{table}[]
\centering
\begin{tabular}{ccccc}
  \toprule
  & $\text{reg}$ 
  & $\text{obj}$
  & $\text{com}$
  & $\text{alt}$ \\ 
  \midrule
  $\text{reg} $ & 6.926 & 6.926 & 5.942 & 7.757 \\ 
  $\text{obj} $ & 6.924 & 6.924 & 5.948 & 7.751 \\ 
  $\text{com }$ & 6.566 & 6.570 & 5.481 & 7.434 \\ 
  $\text{alt} $ & 6.463 & 6.461 & 5.478 & 7.469 \\ 
   \bottomrule
\end{tabular}
\caption{Meta-game for the evolutionary competition between subjective utilities when beliefs are exogenously
  given (see main text).}
\label{tab:ExogeneousEpistemics}
\end{table}
The only evolutionarily stable state of this meta-game is again a monomorphic population of regret types. Accordingly, all
our simulation runs of the (discrete-time) replicator dynamics converge to monomorphic
regret-type populations. The reason why regret-based utilities prosper is because they have a substantial
fitness advantage when paired with imprecise beliefs (Propositions~\ref{proposition1} and \ref{proposition2}). If unmeasurable uncertainty is exogenously given
as something that happens to agents because of the information available in some choice situations, and even if that happens only very infrequently
(i.e., for rather low $p$), regret preferences will outperform objective preferences, as
well as competitive and altruistic preferences.

@@

\iffalse
\subsection{Arbitrary probabilistic beliefs}
\label{sec:arbitr-prob-beli}

So far, we have assumed that epistemic types $\text{prc}$ hold an unbiased, flat belief. It is
worthwhile considering what happens when this assumption is relaxed and we allow agents with
probabilistic beliefs to make use of the full spectrum of probabilistic beliefs. It should be
clear that quality of beliefs directly impacts prospects of successful decision making: if a
decision maker knows the actual state, or can put a high level of credence on the actual state,
accumulated fitness can be expected to be high. We see this, unsurprisingly, in numerical
simulations. We looked at the average payoff accumulated in 100,000 decision problems sampled
as before, except that for belief types with probabilistic beliefs we sampled a random
probabilistic belief $p$ (from an unbiased Dirichlet distribution) and chose the actual world
state with probability weights $p$. This effectively implements a bias for beliefs of the
decision maker that tend to put more weight on the actual state (although the procedure does
admittedly perverse the natural chicken-and-egg logic in this case that the actual world state
should come first and beliefs be a function of it). As a result, the average payoffs of types
$\tuple{\text{reg}, \text{prc}}$ and $\tuple{\text{obj}, \text{prc}}$ are still the exact same
(because we compare what agents with different preference types would do under the same
beliefs; we are not interested in statistical fluctuations given one type better beliefs by
pure happenstance), but with $7.125$ notably higher than before. In effect, better
probabilistic beliefs lead to better decisions. If agents can learn, reason and extrapolate to
form better probabilistic beliefs, that will help them whenever they take their beliefs into
account.

But this is orthogonal to the arguments in this paper, where the focus is on possibilities for
the evolution of non-veridical preferences. We know from Fact~\ref{fact:maxEU-minReg} that
regret types and veridical preference types, when paired with probabilistic belief types that
maximize/minimize expectations, are behaviorally equivalent, \emph{no matter what they believe},
as long as they belief the same. So, if learning, insight and statistical knowledge of a
recurrent situation \emph{can} be brought to bear, this will not make evolution select against
regret-based preferences. If, on the other hand, agents have no basis for probabilistic
beliefs, then we have shown that there are general and basic circumstances in which
regret-based preferences can actually be selected, despite their non-veridicality.
\fi

\section{Related Literature}

do you think we really need a section for this?

@@

Meta-games are useful tools for probing into conceptual questions about ecological
rationality. The approach is rich and general, with many possible further
applications imaginable beyond the case study presented here. For example, the evolutionary
investigation of different risk and ambiguity attitudes when competing against each other may bring new insights on both behavioral and theoretic
decision making under uncertainty, as well as on many issues in theoretical biology and
behavioral ecology. Moreover, variability in the action selection function has been entirely
ignored in this paper. Finally, while this paper only used dummy models of diverse decision
environments, the meta-game approach also allows for empirically informed models of rich and
variable environments.

>>>>

As we noticed, meta-games factor in statistical properties of the environment. For particular
empirical purposes, one could consult a specific class of games $\mathcal{G}$ with appropriate,
maybe empirically informed probabilities $P_{\mathcal{G}}$ in order to match the natural
environment of a given population. For our purposes, let $\mathcal{G}$ be a set of symmetric
two-player games with two acts for a start. A game is then individuated solely by its payoff
function, i.e., by a quadruple of numbers $G=(a,b,c,d)$. As for the occurrence probability
$P_{\mathcal{G}}(G)$ of game $G$, we imagine that the values $a,b,c,d$ are i.i.d.~random
variables sampled from the set $ \lbrace 0, \dots, 10 \rbrace$ according to uniform probability
$P_{V}$. Using Monte Carlo simulations, we can then approximate the values of
equation~(\ref{eq:FittnessChoiceMechGamePairwise}) to construct meta-game payoffs. Results
based on $100,000$ randomly sampled games are given in
table~\ref{tab:ExpectedFitness_4Types}.\footnote{Concretely, $100,000$ games were sampled
  repeatedly by choosing independently four integers between 0 and 10 uniformly at random. For
  each game, the action choices of all four choice mechanisms were determined and payoffs from
  all pairwise encounters recorded. The number in each cell of
  table~\ref{tab:ExpectedFitness_4Types} is the average payoff for the choice mechanism listed
  in the row when matched with the choice mechanism in the column.}

@@

Simulation results of the (discrete time) \emph{replicator dynamics}
\citep{TaylorJonker1978:Evolutionary-St} indeed show that random initial population
configurations are attracted to states with only three player types:
$(\text{reg}, \text{imp})$, $(\text{reg}, \text{prc})$ and $(\text{obj}, \text{prc})$. The
relative proportions of these depend on the initial shares in the population. This variability disappears if
we add a small mutation rate to the dynamics. Take a fixed, small mutation rate $\epsilon$ for
the probability that a player's subjective utility \emph{or} her subjective belief changes to another
utility or belief. The probability that a player's subjective representation randomly mutates into a
completely different representation with altogether different utility and belief would
then be $\epsilon^2$. With these assumptions about ``component-wise mutations,'' numerical
simulations of the (discrete time) \emph{replicator mutator dynamics}
\citep{Nowak2006:Evolutionary-Dy} show that already for very small mutation rates almost all
initial population states converge to a single fixed point in which the majority of players have
regret-based utility. For instance, with $\epsilon = 0.001$, almost all initial populations are
attracted to a final distribution with proportions:

@@

%Obviously, $(\text{reg}, \text{imp})$,
%$(\text{obj}, \text{prc})$, and $(\text{reg}, \text{prc})$ players will be neutrally stable, as before. 
%and fitness-based selection will tend to weed out maxmin play and leave us with a population of
%arbitrary frequency of $(\text{reg}, \text{imp})$, $(\text{obj}, \text{prc})$, and
%$(\text{reg}, \text{prc})$ players. 
The result shows that there is support for the main conceptual
point that we wanted to make: 
%non-veridical representations \emph{can}, under
%specific circumstances, persist under evolutionary selection based on objective
%fitness. Moreover,
objective preference representations are not necessarily favored by natural selection;
objective preferences can be outperformed by non-veridical regret preferences, if agents may
have imprecise beliefs.  This tells us that the main conclusions drawn in the previous section
based on the approximated meta-game of table~\ref{tab:ExpectedFitness_4Types} hold more
generally for arbitrary $2 \times 2$ symmetric games with i.i.d. sampled payoffs.

This result presupposes occasional imprecise beliefs and the maximin decision rule (more on
this below). The assumed imprecise beliefs do not need to be maximally uncertain, however.  Let
the uncertainty held by a player be defined by a closed and convex set of probabilities
$ [s,t] \subseteq \Delta(A) $ over the co-player's actions, where $s$ is the lower probability
and $t$ is the upper probability of action $II$.  We can then prove the following proposition,
which is the equivalent of Proposition~\ref{proposition1} for any possible (not necessarily
maximal) degree of uncertainty $[s,t]$, with $s \neq t$. There is only one difference: we are
now going to require i.i.d. drawing of a \emph{continuous} random variable. This is due to the
fact that, for arbitrarily small intervals $[s,t]$, objective players $([s,t],\text{obj})$ and
regret players $([s,t],\text{reg})$ can behave as holding a unique probability measure (precise
belief) if the underlying payoff space is not dense. The reason for this technical requirement
will become clearer from the proof.

\begin{proposition} \label{proposition2} In the class of $2\times2$ symmetric games generated
  by i.i.d. drawing of a continuous random variable from a distribution with density $P_v$ with
  support $(\underline{r},\overline{r})$, for any imprecise
  belief $[s,t]$, the only evolutionarily stable state of a population with regret players
  $([s,t],\text{reg})$ and objective players $([s,t],\text{obj})$ is a monomorphic state of
  $([s,t],\text{reg})$ players.
\end{proposition}

This tells us that regret-based preferences can also outperform objective preference
representations when agents are capable of learning or otherwise restrict their assumptions
about the co-player's behavior as long as there is, at least on occasion, some imprecision in
their beliefs. We will enlarge on the issue of belief formation after having covered some more
relevant extensions in the next section.

@@

\subsection{More Preference Types}
\label{sec:more-types}

Our claim is not that regret types are the best on the market. We have only compared regret
types to objective types so far. The space of possible preference types is enormous. Let us
look at least at two other types of subjective preferences have been investigated, especially
in behavioral economics and in evolutionary game theory. A famous example is the
\textit{altruistic} preference \citep[e.g.,][]{Beck76,BestGuth98}, summoned to explain the
possibility of altruistic behavior. At the other end of the spectrum, the \emph{competitive}
preference is located. The two subjective utilities are defined as follows:
\begin{enumerate}
\item \textit{altruistic} utility: for all $G \in \mathcal{G}$, $\text{alt}^G(a,a') = \pi^G(a,a') + \pi^G(a',a)$;\footnote{A
    more general formulation would be to define $ \alpha$-altruistic utility, for
    $\alpha \in [0,1]$,
    $ u^G_\alpha(a, a')=\pi^G(a,a') + \alpha \pi^G(a',a)$. Since we are not
    interested in the evolution of degrees of altruism, here we simply fix $ \alpha = 1 $. Analogously for $\alpha$-competitive utilities too.}
\item \textit{competitive} utility: for all $G \in \mathcal{G}$, $\text{com}^G(a,a') = \pi^G(a,a') - \pi^G(a',a)$.
\end{enumerate}

@@

Table~\ref{tab:ExpectedFitness_2x2_Full} shows results of Monte Carlo simulations that
approximate the expected fitness in the relevant meta-game with all the subjective representations considered so
far. These results confirm basic intuitions about altruistic and competitive types: everybody would like to have an altruistic co-player and nobody
would like to play against a competitive player. Perhaps more surprisingly, $(\text{alt}, \text{imp})$
come up strictly dominated by $(\text{com}, \text{imp})$, but competitive types themselves are worse off
against all types except against maxmin players $(\text{obj}, \text{imp})$ than any of
the behaviorally equivalent types $(\text{reg}, \text{imp})$, $(\text{obj}, \text{prc})$, and
$(\text{reg}, \text{prc})$.
\iffalse This is a noteworthy results in the light of the fact
that evolving altruistic preferences have been shown to support cooperative behavior in a
single stage game \myalert{[CITE]}. In contrast, averaging over payoffs in multiple stage
games, like we do here, makes altruistic preferences prime victims of evolutionary eradication.
\fi
It is easy to see then that the previous results still obtain for the larger meta-game in
table~\ref{tab:ExpectedFitness_2x2_Full}: $(\text{reg}, \text{imp})$,
$(\text{obj}, \text{prc})$, and $(\text{reg}, \text{prc})$ are still neutrally stable;
simulation runs of the (discrete-time) replicator dynamics on the $8 \times 8$ meta-game from
table~\ref{tab:ExpectedFitness_2x2_Full} end up in population states consisting of only these
three types in variable proportion.

@@

Results from Section~\ref{sec:results} relied heavily on Fact~\ref{fact:equivalence2x2}, which
is no longer true when we look at arbitrary $n \times n$ games.
Table~\ref{tab:ExpectedFitness_10x10} gives approximations of expected fitness in the class of
$n \times n$ symmetric games. Concretely, the numbers in table~\ref{tab:ExpectedFitness_10x10}
are averages of evolutionary payoffs obtained in 100,000 randomly sampled symmetric games,
where each game was sampled by first picking a number of acts $n \in \set{2, \dots, 10}$
uniformly at random, and then filling the necessary $n \times n$ payoff matrix with
i.i.d.~sampled numbers, as before.

@@

On the other hand, there are plenty of ways in which the basic insight from
Propositions~\ref{proposition1} and \ref{proposition2} can make for situations in which
evolution would favor regret types, even in $n \times n$ games. If, for example, the belief of
a player is a trait that biological evolution has no bite on, but rather something that the
particular choice situation would exogenously give us (possibly because of the different amount
of information available in different choice situations), then regret-based preferences can
again drive out veridical preferences altogether. For example, suppose that only preferences
compete and that agents' beliefs are exogenously given, in such a way that both players hold
precise (Bayesian) uniform beliefs with probability $p$ and they both have maximally imprecise
beliefs otherwise. This transforms the meta-game from table~\ref{tab:ExpectedFitness_10x10}
into a simpler $4 \times 4$ meta-game in which the payoff obtained by a subjective utility is
the weighted average over the payoffs of the subjective representations including that utility
in table~\ref{tab:ExpectedFitness_10x10}. Setting $p = 0.98$ for illustration, we get the
meta-game in table~\ref{tab:ExogeneousEpistemics}.
\begin{table}[]
\centering
\begin{tabular}{ccccc}
  \toprule
  & $\text{reg}$ 
  & $\text{obj}$
  & $\text{com}$
  & $\text{alt}$ \\ 
  \midrule
  $\text{reg} $ & 6.926 & 6.926 & 5.942 & 7.757 \\ 
  $\text{obj} $ & 6.924 & 6.924 & 5.948 & 7.751 \\ 
  $\text{com }$ & 6.566 & 6.570 & 5.481 & 7.434 \\ 
  $\text{alt} $ & 6.463 & 6.461 & 5.478 & 7.469 \\ 
   \bottomrule
\end{tabular}
\caption{Meta-game for the evolutionary competition between subjective utilities when beliefs are exogenously
  given (see main text).}
\label{tab:ExogeneousEpistemics}
\end{table}
The only evolutionarily stable state of this meta-game is again a monomorphic population of regret types. Accordingly, all
our simulation runs of the (discrete-time) replicator dynamics converge to monomorphic
regret-type populations. The reason why regret-based utilities prosper is because they have a substantial
fitness advantage when paired with imprecise beliefs (Propositions~\ref{proposition1} and \ref{proposition2}). If unmeasurable uncertainty is exogenously given
as something that happens to agents because of the information available in some choice situations, and even if that happens only very infrequently
(i.e., for rather low $p$), regret preferences will outperform objective preferences, as
well as competitive and altruistic preferences.

@@

\subsection{Informed or Sophisticated Beliefs}

Since our main purpose here was to illustrate the potential usefulness of a meta-game approach
by a comparison of regret and objective preference schemes, we have neglected entirely a very
important and interesting issue, namely the evolution of ways of forming more concrete beliefs
about a co-player's likely behavior or the actual state of the world. It is obvious (and shows
in numerical simulations) that agents whose beliefs are more likely to put more weight on the
actual world state (in solitary decisions) or on the co-player's actual choice will accrue
higher fitness-payoffs than radically uncertain players. For reasons of space we must,
unfortunately, leave a deeper exploration of belief type evolution to another occasion. Two
remarks are in order nonetheless. Firstly, belief type evolution can be studied without
conceptual hurdles in a meta-game approach, so that there is no principled argument here
against the main methodological contribution of this paper. Secondly, our results regarding the
comparison between regret and objective preference types remain to be informative, even if we
allow agents to learn or reason strategically. This is because we know from
Fact~\ref{fact:maxEU-minReg} that regret and objective preferences come up behaviorally
equivalent when paired with precise probabilistic beliefs (given a maximin decision rule). This
holds no matter what the content of that belief is. So, if learning, reasoning or statistical
knowledge of a recurrent situation \emph{can} be brought to bear, this will not make evolution
select against regret-based preferences. If, on the other hand, agents resort to imprecise
beliefs at least occasionally (e.g., when they are unaware of the co-player or her utilities,
or when strategic reasoning cannot reduce all uncertainty about the opponent's choice), then we
have shown that there are general and basic circumstances in which regret-based preferences
could even be favored by natural selected over objective preferences.

@@

Meta-games are useful tools for probing into conceptual questions about ecological
rationality. The approach is rich and general, with many possible further applications
imaginable beyond the case study presented here. For example, the evolutionary investigation of
different risk and ambiguity attitudes when competing against each other may bring new insights
on both behavioral and theoretic decision making under uncertainty, as well as on many issues
in theoretical biology and behavioral ecology. Moreover, this paper only considered one
decision rule (maximin), leaving it to future work to explore evolutionary competition here as
well. Finally, while this paper only used dummy models of diverse decision environments, the
meta-game approach also allows for empirically informed models of rich and variable
environments.
